---
title: "FIFA 19 Final Write-Up"
author: "SPT Squad"
date: "December 13, 2019"
output: github_document
---

<!-- ====================== Packages and data cleaning ===================== -->

```{r suppress-code, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)
```

<!-- Load packages -->
```{r load-packages}
library(broom)
library(dplyr)
library(readr)
library(tidyverse)
library(infer)
library(knitr)
```

<!-- Load, filter, and clean data -->
```{r load-and-clean}
rawdata <- read_csv("../data/data.csv")

# Remove NA values
rawdata <- na.omit(rawdata)

# Pull a random sample of 1000 players.
# This is for ease of analysis while retaining predictive/generalizing power.
set.seed(5)
fifa <- sample_n(rawdata, 1000)

# Remove currency signs
fifa$Value <- substring(fifa$Value, 2)
fifa$Wage <- substring(fifa$Wage, 2)

# Convert currency to numeric type
fifa <- fifa %>%
  mutate(Wage = case_when(
    str_sub(Wage, -1) == "M" ~ (parse_number(Wage)*1000000),
    str_sub(Wage, -1) == "K" ~ (parse_number(Wage)*1000)
  ), Value = case_when(
    str_sub(Value, -1) == "M" ~ (parse_number(Value)*1000000),
    str_sub(Value, -1) == "K" ~ (parse_number(Value)*1000)
  ))

# Refactor `International Reputation`
fifa <- fifa %>%
  mutate(`International Reputation` = case_when(
    `International Reputation` == 1 ~ "Unknown",
    `International Reputation` == 2 ~ "Coming Up",
    `International Reputation` == 3 ~ "Well-Known",
    `International Reputation` == 4 ~ "Star",
    `International Reputation` == 5 ~ "Superstar"
  )) %>%
  mutate(`International Reputation`= factor(`International Reputation`,
                                            levels = c("Unknown", 
                                                       "Coming Up",
                                                       "Well-Known", 
                                                       "Star", 
                                                       "Superstar")))
```

<!-- ============================== Write-Up ============================== -->
# Introduction
<!-- Is the research question, motivation for the analysis, and data clearly 
described? -->

FIFA, also known as FIFA Soccer or FIFA Football, is a series of soccer video game simulations released annually by EA Sports. The FIFA series is listed in the Guinness World Records as the best-selling sports video game franchise in the world, by 2018, the FIFA series had sold over 260 million copies.

Among the FIFA community, player ratings are a highly debated topic. A player’s overall rating is the most universal way to measure how good a player is in FIFA. Soccer players in real life often care about these ratings too!

Should Lionel Messi deserve to have a higher rating than Cristiano Ronaldo? Is Neymar rated too highly given that he is a young player? Are offensive players worth more in the transfer market compared to defensive players? Through our final project, we perform statistical inference on the characteristics related to a high player rating, as well as compare players between different groups (position, foot dominance, etc). Our specific research questions are identified later in this analysis.

## Dataset

Our particular dataset focuses on the iteration of the game from 2019, FIFA 19. The dataset is from [Kaggle](https://www.kaggle.com/karangadiya/fifa19). It includes detailed attributes for every registeded player in the game (18,000+). The data was scraped from a comprehensive online database for FIFA enthusiasts, https://sofifa.com/. Some of the more important variables include: full name, age, nationality, overall rating (out of 99), club, wage, dominant foot, international reputation, body type, position, among many other attributes across players. The benefit of the dataset is that it is all-encompassing, it has the attributes for every single player in the game so we are working with a random sample of the entire population of FIFA players in FIFA 19 (no sampling or response bias), which means our inference and analysis will be applicable to the entire population of FIFA players. Additionally, the dataset is very comprehensive: there are a lot of variables to work with which will give our model stronger predictive power as well as the ability to fine-tune the independent variables.

## Univariate Analysis
<!-- Are univariate distributions in the data clearly visualized and described 
using the appropriate summary statistics? -->

Before we explore relationships between our variables, we visualize and look at the summary statistics for some of our important variables: `Overall`, `Potential`, `Age`, `Preferred Foot`, and `International Reputation`.

#### Overall Player Rating
```{r univ-overall-dist}
ggplot(fifa, mapping = aes(x = Overall)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Distribution of Overall Player Ratings", 
       x = "Overall Rating (From 0 to 99)", y = "Frequency")
```

```{r summarise-overall, results='hide'}
summarise(fifa, mean = mean(Overall), sd=sd(Overall)) 
```
The distribution of `Overall` is approximately normal, with a mean of 66.02 and standard deviation of 7.01. This is to be expected -- the game must have a balance of players with low, medium, and high overall ratings.

#### Player Potential

```{r univ-potential-dist}
ggplot(fifa, mapping = aes(x = Potential)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Distribution of Potential Player Ratings", 
       x = "Potential Rating (From 0 to 99)", y = "Frequency")
```

```{r potential-summary-stats, results='hide'}
summarise(fifa, mean = mean(Potential), sd=sd(Potential)) 
```

The distribution of `Potential` is also approximately normal; its mean is 70.96 and its standard deviation is 6.14. It is worth nothing that the mean of `Potential` is higher than the mean of `Overall.` This makes sense to us -- many young players have lower `Overall` ratings but higher `Potential` ratings, relatively.

#### Player Age
```{r univ-age-dist}
ggplot(fifa, mapping = aes(x = Age)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Distribution of Player Age", 
       x = "Current Age", y = "Frequency") + geom_density()
```

```{r age-summary-stats, results='hide'}
summarise(fifa, mean = mean(Age), sd=sd(Age))
```

The distribution of player ages is relatively symmetric, with mean 25.3 and standard deviation 4.84. There are some very young players in our dataset (younger than 18) and some older players (older than 35).

#### Player Preferred Foot

```{r univ-pref-foot-dist}
ggplot(fifa, mapping = aes(x = `Preferred Foot`)) +
  geom_bar() +
  labs(title = "Frequency of Player Footedness", 
       x = "Preferred Foot", y = "Frequency")
```

```{r pref-foot-prop, results='hide'}
fifa %>%
  count(`Preferred Foot`) %>%
  mutate(proportion = n/sum(n))
```

Unsurprisingly, the large majority of players prefer their right foot. Specifically, 22.6% of players in our random sample prefer their left foot while 77.4% of players in our random sample prefer their right foot.

#### Player International Reputation

```{r univ-intl-rep}
ggplot(fifa, mapping = aes(x = `International Reputation`)) +
  geom_bar() +
  labs(title = "Frequency of International Reputation", 
       x = "International Reputation", y = "Frequency")
```

```{r intl-rep-prop, results='hide'}
fifa %>%
  count(`International Reputation`) %>%
  mutate(proportion = n/sum(n))
```

From our visualization and a calculated relative frequencies, we can see that an overwhelming majority of the players in our random sample (91.6%) have the `International Reputation` designation of “Unknown”. Only 0.4% of players in our random sample have “Star” as their designation for `International Reputation`. Our random sample of 1000 players does not yield any players that have the `International Reputation` value of “Superstar” -- this isn't too surprising to us, as there are only 6 players in the entire dataset who have this designation.



## Bivariate/Multivariate Analysis
<!-- Are relevant and interesting bivariate/multivariate relationships 
clearly visualized and described using the appropriate summary statistics? -->

We'll explore the following variables and relationships through our bivariate analysis:  `Value` vs. `Wage`, `Overall` vs. `Finishing`, `Value` vs. `Finishing`, `Potential` vs. `Position`, `Overall` vs. `Preferred Foot`, and `Overall` vs. `International Reputation`.

#### Market Value and Player Compensation

```{r value-vs-wage}
ggplot(fifa, aes(x = Wage, y = Value)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(x = "Player Wage (in Euros)",
       y = "Player Market Value (in Euros)", 
       title = "Relationship between Wage and Market Value") +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma)
```

Wage is the amount of money that a club will have to pay for a certain player in FIFA. The market value is the transfer value in which a club must pay to acquire a player. The amount of money that a player earns in FIFA should be related to their transfer market value. We observe a positive relationship -- the higher a player’s wage, the higher their market value. We see that the vast majority of the data is concentrated in the bottom left of the graph, with a few players having much higher wages and market values.

#### Finishing Ability and Overall Player Rating

```{r overall-vs-finishing}
ggplot(fifa, aes(x = Finishing, y = Overall)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(x = "Finishing Ability Rating (From 0 to 99)",
       y = "Overall Rating (From 0 to 99)", 
       title = "Relationship between Finishing Ability and Overall Rating")
```

The ability for a soccer player in FIFA (and in real life) to finish around the net is crucial. Therefore, we look at the relationship between players’ finishing ability (rated from 0 to 99) and their overall rating (rated from 0 to 99). As expected, we see a positive relationship between finishing ability and overall rating.

#### Finishing Ability and Market Value

```{r finishing-vs-market}
ggplot(fifa, aes(x = Finishing, y = Value)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  labs(x = "Finishing Ability Rating (From 0 to 99)",
       y = "Market Value (in Euros)", 
       title = "Relationship between Finishing Ability and Market Value") +
  scale_y_continuous(labels = scales::comma)
```

Now we look at the relationship between finishing ability and market value. There is a positive relationship between finishing ability and market value - those who have higher finishing ability ratings are expected to have higher market values. It is also worth noting that at the higher end of finishing ability ratings some players make a lot more than others with similar finishing ratings. This could be because of other factors such as `International Reputation` or even the team that the player plays for.

#### Position and Potential Rating

```{r potential-vs-position}
ggplot(data = fifa, mapping = aes(x = `Position`, y = `Potential`)) +
  geom_boxplot() +
  labs(x = "Position",
       y = "Potential Rating (From 0 to 99)", 
       title = "Potential Rating by Position")
```

In this visualization, we use a boxplot to look at players’ potential rating (from 0 to 99). The potential rating is important to see which players have a chance to increase their current overall rating. We see here that players that play the left field position generally have higher potential ratings compared to other positions.

#### Foot Preference and Overall Rating

```{r overall-vs-foot}
ggplot(data = fifa, mapping = aes(x = `Preferred Foot`, y = `Overall`)) +
  geom_boxplot() + 
  labs(x = "Preferred Foot",
       y = "Overall Rating (From 0 to 99)", 
       title = "Overall Rating by Foot Preference")
```

Many of the most famous and talented FIFA players are left-footed, such as Lionel Messi and Arjen Robben. In this visualization, we want to compare overall ratings by foot preference. The visualization shows that left-footed and right-footed players are very similar when it comes to overall rating.

#### International Reputation and Overall Rating

```{r overall-vs-intl}
ggplot(data = fifa, mapping = aes(x = `International Reputation`,
                                  y = `Overall`)) +
  geom_boxplot() +
  labs(x = "International Reputation",
       y = "Overall Rating (From 0 to 99)", 
       title = "Overall Rating by International Reputation")
```

In this visualizaiton we explore overall rating by international reputation. The results are what we expected - those who have “Star” designation have the highest overall rating.

With these findings from our univariate and bivariate analyses, we're ready to explore our research questions.

# Research Questions, Data Analysis & Results
<!-- To include from data analysis write-up -->

```{r load-data, message = FALSE}
fifa <- read_csv("../data/dataWorkingSample.csv")
```

## Question One: What are the characteristics that are important in determining a player's overall rating?

**Dependent variable:** `Overall` (overall rating)  
**Independent variables:** `Age`, `Value`, `Wage`, `International Reputation`, `Potenial`, `Finishing`, `Preferred Foot`, `Position`  
**Comparison groups:** `Nationality`, `Club`, `Body Type`  
**Hypothesis:** Out of the above independent variables, we expect `Value`, `Wage`, `International Reputation`, `Potential`, and `Finishing` to have a statistically significant impact on overall rating based on our exploratory data analysis.  

#### Which method?
We use backwards selection with AIC as our decision criteria in order to determine our selected model. We selected this statistical method to answer this question because backwards selection is a good method to narrow down the appropriate variables for best fitting the data based on a full model.  

#### Recoding `Position`
Before we fit our models, we'll start with a little bit of data cleaning. Specifically, we want to re-group individual positions as either `offense` or `defense` to simplify our analysis:

```{r newvariables}
fifa_new <- fifa %>%
  mutate(Position = case_when(
    Position %in% c("ST", "LS", "RS", "LM", "RM", "LW", "CF", "RW", "LF", "RF", 
                    "CDM", "LDM", "RDM", "CM", "LCM", "RCM", "CAM", "LAM", 
                    "RAM") ~ "offense",
    Position %in% c("CB", "RCB", "LCB", "RB", "LB", "GK", "LW", "RW", "LWB", 
                    "RWB") ~ "defense"
  ))
```

#### Creating our full model
Our full model is the regression of `Overall` on the independent variables that we want to investigate. Out of our curiosity, we also include three interaction terms in the full model: `Age*Value`, `International Reputation*Finishing`, and `Value*International Reputation`.

```{r question1-full-model}
tidy((m_full <- lm(Overall ~ Age + Value + Wage + 
                     `International Reputation` + Potential + Finishing + 
                      `Preferred Foot` + Position + Age*Value + 
                     `International Reputation`*Finishing +
                     Value*`International Reputation`,
                   data = fifa_new)))
```

#### Finding our selected model with backwards selection

From this full model, we'll run backwards selection:

```{r q1-backwards-selection}
selected_model <- step(m_full, direction = "backward")
```

#### Primary conclusions

Based on the results, our hypothesis was partially correct. Specifically, we were correct about `Value`, `Wage`, `International Reputation`, `Potential`, and `Finishing` being variables that would be significant; however, there was one variable that we did not think would be significant that ended up being significant, `Age`. 

The variables that ended up in the selected final model are: `Age`, `Value`, `International Reputation`, `Potential`, `Finishing`, `Position`, and `Value*International Reputation`. Therefore, out of the variables that we were considering at the onset of our analysis (included in the full model), these final variables are significant in determining a player's overall rating. Out of the three interaction effects we explored in the full model, `Value*International Reputation` was included in the selected model.

The p-values in the selected final model are also worth noting:

```{r q1-p-values}
tidy(selected_model) %>% 
  select(term, estimate, p.value) %>% 
  kable(format = "markdown", digits = 7)
```

Specifically, a variable like `International Reputation` appears much more subjective and less rigid in determining -- we can see that there are large fluctuations in the p-values for the different factor levels. Nonetheless, `International Reputation` still proves an important characteristic in determining Overall Rating by our AIC backwards model selection.

#### Visualizations

We might be able to better understand the influence of `International Reputation` by visualizing it against overall rating:

```{r q1-intlrep-overall}
fifa_new %>%
  ggplot(aes(x = `International Reputation`, y = Overall, color = Position)) +
  geom_boxplot() +
  labs(x = "International Reputation",
       y = "Player Overall Rating (From 0 to 99)", 
       title = "Distribution of Overall Rating by International Reputation",
       subtitle = "Colored by Position (`defense` or `offense`)") +
  scale_y_continuous(labels = scales::comma)
```

This boxplot shows us the distributions of players' overall ratings by category of international reputation. To reflect the data cleaning we started with, and in line with our second research question, we further grouped these plots by `Position`, with red/pink plots reflecting ratings for defensive players and teal/blue plots reflecting those for offensive players.  

The trends in center across reputation groups is exactly as we'd expect: highest for international stars, followed by well-known players, up-and-coming players, and lastly unknown players. Within International Reputation categories, overall ratings do not seem to vary much between defensive and offensive positions. The spread of overall ratings is tightest for Star players; this might be because there are fewer players in that category, or else because the requirements for a player to be considered a Star are much more stringent than for the other categories. Unknown players have the widest spread; they also are the only category that reflects outliers in this box plot. This makes sense to us: As not much is known about them on the global stage, unknown players can be wild cards with a wide range of ability and performance. We would expect there to be some fantastic players overlooked by the global media (perhaps playing on mediocre teams, or from a smaller media market). We would also expect players with a low overall rating to be unknown internationally -- with the worst performance, there's no reason for them to be famous.

Let's also take a look at how age and overall rating are related:

```{r q1-vis-age-overall}
fifa_new %>%
  ggplot(aes(x = Age, y = Overall)) +
  geom_point() +
  geom_smooth(method = "auto", se = FALSE, colour = "red") +
  labs(x = "Player Age",
       y = "Player Overall Rating (From 0 to 99)", 
       title = "Relationship between Age and Overall Rating") +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma)
```

We used `geom_smooth()`'s `"auto"` method type to visualize this trend, as we didn't expect there to be a strictly linear relationship. This proved accurate: As we can see from the red line, there's a mostly-linear, positive relationship between age and overall rating to a point, then rating appears to level off and slightly decrease as players age. A player's rating peak seems to hit at around age 27, after which age has a slightly negative effect on overall rating. This makes sense to us! In general, younger players lack the experience, consistency, and legacy to earn them higher overall ratings. Similarly, as injuries and the effects of age accumulate, we would expect the oldest players to have slightly lower ratings. Lastly, the lowest overall ratings are mostly consolidated among younger players. We believe this to be because poor-performing players would leave the league before lasting long enough to reach comparatively old age, while better players stay in the game longer.

## Question Two: Do players who are certain positions have higher market value compared to other positions? More specifically, do offense players have higher market value compared to their counterparts who play defense? 

**Dependent variable:** `Value` (market value)  
**Independent variables:** `Position` (all associated dummy variables because `Position` is a factor variable)  
**Comparison group:** `Club`, `Nationality`, `Preferred Foot`  
**Hypothesis:** If a player's postition is an offense position, then he will be traded at a higher market value in comparison to an otherwise similar player whose position is defensive. This makes sense - the most heralded players that are paid the most tend to be that score the most goals (usually offensive players); however, there are some very high profile defenders and in FIFA as well.

To understand differences across groups -- here, `offense` and `defense` positions -- we'll use hypothesis tests. We've already regrouped the given positions as offense or defense; our code for this is above, under research question 1. With that out of the way, let's take a look at how a player's market value is distributed by position type.

#### Understanding market value

We'll start with summary statistics. First, among offensive players:
```{r off-summary-statistics}
options(scipen = 999)

#Offense
fifa_new %>%
  filter(Position == "offense") %>%
  summarize(`Mean` = mean(Value), 
            `Median` = median(Value),
            `Standard Deviation` = sd(Value), 
            `n` = n())
```

And next, among defensive players:
```{r def-summary-statistics}
#Defense
fifa_new %>%
  filter(Position == "defense") %>%
  summarize(`Mean` = mean(Value), 
            `Median` = median(Value),
            `Standard Deviation` = sd(Value), 
            `n` = n())
```

We'll continue with a visualization of value distribution by position:

```{r value-plot, message = FALSE}
ggplot(fifa_new, aes(x = Value)) +
  geom_histogram() +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  facet_grid( Position ~ .) +
  theme(text = element_text(size= 10),
        axis.text.x = element_text(angle = 50, hjust = 1)) +
  labs(title = "Value distribution of players by their position", 
       x = "Value", 
       y = "Number of Players")

```

With players' values heavily right-skewed, this is a difficult visualization to understand. Let's try again, this time visualizing the distribution with `log(Value)`: 

```{r q2-log-value-vis, message = FALSE}
ggplot(fifa_new, aes(x = log(Value))) +
  geom_histogram() +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  facet_grid( Position ~ .) +
  theme(text = element_text(size= 10),
        axis.text.x = element_text(angle = 50, hjust = 1)) +
  labs(title = "Value distribution of players by their position", 
       subtitle = "Player count by log of `Value`",
       x = "Log of Value", 
       y = "Number of Players")
```

This visualization shows us a bit more information -- specifically, it looks like offensive players have generally higher values. We can confirm this with a hypothesis test. Since the distributions of Value for both offensive and defensive players are skewed, we'll compare the medians.

#### Median Value of Offense > Median Value of Defense?
**Hypothesis:** Let *median* represent the median Value of players:

H0: *median*Offense - *median*Defense = 0

Ha: *median*Offense - *median*Defense > 0

**Calculate the observed sample statistic:**
```{r offense-defense_obs}
sampmed_diff <- fifa_new %>%
  specify(Value ~ Position) %>%
  calculate(stat = "diff in medians", order = c("offense", "defense"))

sampmed_diff
```

Using the difference in median value, we have a sample statistic value of `r sampmed_diff`.  

We'll continue by generating and visualizing the null distribution. As the research question asks for a difference between, we're looking at an independence hypothesis. This has us using `"permute"` in our generate function.

**Permute the null distribution:**
```{r offense-defense_permute}
off_def_perm <- fifa_new %>%
  filter(Position %in% c("offense", "defense")) %>%
  specify(response = Value, explanatory = Position) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in medians", order = c("offense", "defense"))
```

Let's visualize the distribution we generated:

```{r offense-defense_visualise}
visualize(off_def_perm) +
  labs(title = "Null distribution of differences in medians",
       subtitle = "Difference in sample median, Offense - Defense",
       x = "sampmed_Offense - sampmed_defense",
       y = "Count") +
 shade_p_value(obs_stat = sampmed_diff, direction = "greater")
```

We shaded the p-value using the "greater" direction, as our calculation was taken in the order `Offense` - `Defense`, and we're looking to see if offensive players have a higher market value than defensive players. Let's calculate the p-value to confirm:

**Find the p-value:**
```{r offense-defense_p}
get_p_value(off_def_perm, sampmed_diff, direction = "greater")
```

#### Primary conclusion
With a p-value smaller than 0.05, we reject the null hypothesis in favor of the alternative hypothesis. The data provides convincing evidence that the median market value of offensive players is greater than the median market value of defensive players.

Our hypothesis was indeed correct - we hypothesized that players who play offense would have a higher transfer `Value` when compared to players who play defense. As discussed before, our intuition was that the players who score the most goals (typically offense players) should be worth the most in the transfer market. We did not hypothesize how much higher offense player `Value` would be compared to defense player `Value`. In the following portion of this analysis, we will construct a confidence interval to better understand the difference in median `Value` for offense and defense players. 

#### Confidence Interval
First, we'll generate a bootstrap distribution. Then using this distribution, we calculate our confidence interval:

```{r off_def_boot}
bootstrap_distribution <- fifa_new %>%
  specify(response = Value, explanatory = Position) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "diff in medians", order = c("offense", "defense"))
```

```{r off_def_conf_calc}
ci_bounds <- get_ci(bootstrap_distribution, level = 0.95)
ci_bounds
```

With this calculation, we are 95% confident that the median `Value` for offensive players is between 75,000 and 275,000 Euros higher than that for defensive players.

Let's see what this confidence interval looks like in the context of our bootstrap distribution:

```{r off_def_conf_vis}
visualise(bootstrap_distribution) +
  shade_confidence_interval(endpoints = ci_bounds)
```

Our 95% confidence range is shaded above in teal, with our lower bound (75000) and our upper bound (275,000) marked with the solid green lines.

#### Putting it all together
Building on what we learned from our earlier visualizations, we'll use faceting to better understand the relationship between market value and position. With this next visualization, we'll add the context of international reputation and preferred foot:

```{r visual-q2, fig.width = 10}
fifa_new %>%
  ggplot(aes(x = Position, y = Value, color = `Preferred Foot`)) +
  geom_boxplot() + 
  facet_grid(. ~ `International Reputation`) + 
  labs(title = "FIFA Player Value",
       subtitle = "Faceted by International Reputation, colored by Preferred Foot",
       x = "Position, within level of International Reputation", 
       y = "Player Market Value (in Euros)")
```

There's a lot happening in this visualization. Players are grouped initially by category of international reputation (as we saw in our first research question). Within these faceted groups, players are next grouped by defensive or offensive position. Within these position groups, the box plots are colored by preferred foot.  

Some observations we found interesting: None of the star players included in our dataset prefer to play with their left foot. As in our earlier visualizations of international reputation, outliers are much more heavily present with the `Unknown` category. `Well-Known` and `Star` players have the widest ranges in market value. For `Coming-Up` players, looking at the trend in centers shows us that right-footed players seem to have a slightly higher market value. This is different among `Well-Known` players; these left-footed players (all offensive) have a much higher market value than offensive players, by both the center and spread of the distribution. Player market value by international rating appears to generally follow the same ordering we saw in overall rating above: Star, Well-Known, Coming-Up, and Unknown, though the high density of outliers among Unknown complicates that generalization relative to Coming-Up players.

# Conclusion
<!-- One to two pages -->

## Research Questions
<!-- for each, a summary on what we've learned about our research question and statistical arguments supporting our conclusions. Start with EDA!-->

#### Question One: What are the characteristics that are important in determining a player's overall rating?

The exploratory data analysis was a crucial step of the analysis and construction of our methodology in tackling Question One. We analyzed a vareity of different visualizations to narrow down which variables we want to include as part of our full model to be narrowed down through backwards selection. We learned as a group that understanding the dataset that you are working with and performing a thorough exploratory data analysis before jumping into inference is good practice. 

As we have noted in the above section, the variables that ended up in the selected final model are: `Age`, `Value`, `International Reputation`, `Potential`, `Finishing`, `Position`, and `Value`X`International Reputation`. Now we have a much better idea what causes a FIFA player to have a certain rating, and we can even predict what a FIFA player's overall rating might be given values for these independent variables! 

Let's evaluate the predictive power of our selected model using R-Squared and see how backwards selection performed:

```{r rsquared}
glance(selected_model)$r.squared
```

The R-Squared is .8908, which means that roughly 89.08% of the variability in overall rating can be explained the covariates in our selected model. This is quite high - which is a result of careful initial data analysis, initial variable selection for the full model, as well as the ability of the backwards selection method to select the best-fitting model using the variables from the full model. Backwards model selection is a useful tool to use when determining apprporiate regression models and narrowing down a list of independent variables. Some limitations of this method will be discussed in the Criqitues and Limitations portion of the conclusion. Nonetheless, we will definitely continue to use this tool in the rest of our data science careers. 

#### Question Two: Do players who are certain positions have higher market value compared to other positions? More specifically, do offense players have higher market value compared to their counterparts who play defense? 

Again, our exploratory data analysis pointed us in the right direction to flesh out this research question. Specifically, we observed that the distribution of Value was heavily skewed; therefore, instead of using the CLT and analyzing the mean Values for offensive and defensive players, we analyzed the medians. If we had not looked at these visualizations and used the group means instead, our results would be biased. Skipping an initial data analysis will negatively impact the rest of one's analysis!

As stated in the above section, we determined that offensive players do indeed have a higher Value in the trasfer market when compared to defensive players. We then went on to construct a confidence interval that says that we are 95% confident that the median Value for offensive players is between 75,000 and 275,000 Euros higher than that for defensive players.

In constructing and researching confidence intervals, we also learned an important caveat in terms of their interpretation. We are careful to say that we are "95% confident" instead of there being a "95% probability". This is because we have already observed the data after the confidence interval is constructed, there is nothing random in our interval. We had wondered why it is correct to say "95% confident" instead of "95% probability" and now we have much more clarity about it. 

## Critiques, Limitations, and Future Work
<!-- critique your own methods and provide suggestions for improving your analysis. Issues pertaining to the reliability and validity of your data and appropriateness of the statistical analysis should also be discussed here. include a paragraph on what you would do differently if you were able to start over with the project or what you would do next if you were going to continue work on the project -->

For the ease of the project and due to the difference in our group members' computers' computing powers, we decided to work with a random sample of players from the population of players of size 1000. This project is a very rare example of a time when statisticians have data for an entire population to perform statistical analysis on - and the most robust and correct form of analysis would feature analysis performed on the entire population. While a sample size of 1000 is more than large enough to perform sound statistical analysis, all of our tests would be more robust with the entire population (or even a larger sample size). Additionally, a downfall of random sampling is that there is no guarantee that the data is reflective of the population on average. There were 6 very high-profile players that were given the International Superstar designation in the `International Reputation` variable; however, our simple random sample did not capture any of them (which makes sense but we were interested in seeing how these high-profile players would impact our analysis). On the other hand, because we did have the entire population at our disposal and took a random sample, there was no selection bias. 

As highlighted above, for Question One, our selected model has quite strong predictive power with an R-Squared value of .8908. However, the R-Squared value does not determine whether the coefficient estimates and predictions are biased, so it is always good practice to look at residual plots. It is possible for a biased model to have high R-Squared! 

Let's take a look at a residuals plot against one of our independent variables, `Age`:

```{r residplot}
selected_model.resid <- resid(selected_model)
fifa_new %>%
  ggplot(aes(x = Age, y = selected_model.resid)) +
  geom_point() + geom_hline(yintercept = 0) +
  labs(x = "Player Age",
       y = "Residuals", 
       title = "Relationship between Age and Overall Rating")
```

An unbiased model has residuals that are randomly scattered around 0. Our residuals are not randomly scattered around 0, but rather follow some sort of pattern - this is not good and a limitation of our analysis. One of the assumptions of a least-squared regression is homoscedastic residuals (residuals are drawn from a population with constant variance); however, as we can see from our residuals plot our residuals are heteroscedastic, which signals that our model may indeed be biased. Additionally, this is only the residual plot for the residuals against one of our independent variables! 

If we were to continue working with this project, we would indeed analyze the residual plots for all our independent variables and then go about rectifying the possible heteroscedasticy either through building and testing different models (potentially median regression) or testing different transformations such as log transformations. While backwards model selection is a good method to find models with low AIC and high R-Squared, there are clearly limitations to this method. Using solely R-Squared to determine how good a model is is not good statistical practice. Selecting and fine-tuning new regression models would be a worthwhile next step for our project.